The finite representation property for some reducts of relation algebras

### 2021-01-25 frp-paper/Paper.tex
wip

\subsection{Axiomatisating ${\bf R}(+, ;)$ and the finite representation property}

@@

\end{itemize}
Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By elementary chain theorem (see \cite[Theorem 2.5.2]{hodges1997shorter})

@@

@@

  \begin{enumerate}
    \item $(\Rightarrow)$
    \item $(\Leftarrow)$
  \end{enumerate}
\end{proof}

Now we are seeking to obtain the explicit axiomatisation of ${\bf R}(;, +)$. The idea is to axiomatise this class with the additional axioms claiming that $\exists$ has a winning strategy in certain games. See \cite[Chapter 9]{hirsch2002relation} for the discussion in detail to have a more general methodology. Our purpose is to provide the recursively enumerable axiomatisation of ${\bf R}(;, +)$. Thus, according to famous Craig's trick \cite{craig1953axiomatizability}, ${\bf R}(;, +)$ will have the decidable axiomatisation.

\begin{definition}

\end{definition}

@@

Let $\mathcal{N} = \langle E, V, l \rangle$ be a term prenetwork, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation. Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with the labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l(x, y)\}$.

@@

  Let $\mathcal{N}$ be a term network.

  ...

@@

    \item $\sigma_{n + 1} = \sigma_{n} \land $
  \end{enumerate}
\end{proof}

\begin{theorem}
  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$ defined in the proof of Lemma~\ref{ax}.
\end{theorem}

@@

\begin{corollary}
  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
\end{corollary}

>>>>

\subsection{Axiomatisating ${\bf R}(+, ;)$}

@@

\end{itemize}
Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By Tarksi-Vaught theorem on elementary chain union (see \cite[Theorem 2.5.2]{hodges1997shorter})

@@

The following proposition is a version of \cite[Proposition 7.24]{hirsch2002relation} and the right-to-left part is proved using K\"{o}nig's lemma \cite[Exercise 5.6.3]{hodges1997shorter}.

@@

Now we are seeking to obtain the explicit axiomatisation of ${\bf R}(;, +)$. See \cite[Chapter 9]{hirsch2002relation} for the discussion in detail to have a more general methodology. Our purpose is to provide the recursively enumerable axiomatisation of ${\bf R}(;, +)$. Thus, according to famous Craig's trick \cite{craig1953axiomatizability}, ${\bf R}(;, +)$ will have the decidable axiomatisation. According to the proposition above, it is sufficient to extract the axioms claiming that $\exists$ has a winning strategy in all finitary games.

@@

Let $\mathcal{N} = \langle E, V, l \rangle$ be a term prenetwork, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation. Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l_{\mathcal{N}}(x, y)\}$.

@@

  Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ that reflect the composition, witness, and join moves respectively:
  \begin{enumerate}
    \item $\mathcal{N}_c = $
    \item $\mathcal{N}_w = $
    \item $\mathcal{N}_j = $
  \end{enumerate}
  ...

@@

This formula merely claims that $\exists$ has a winning strategy in the zero length game.

    \item $\sigma_{n + 1}(\mathcal{N}, v) = \\ \sigma_{n}(\mathcal{N}_c, ..) \land \sigma_{n}(\mathcal{N}_w, ..) \land \sigma_{n}(\mathcal{N}_j, ..)$
  \end{enumerate}
\end{proof}

\begin{theorem}
  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$ defined in the proof of Lemma~\ref{ax}. Moreover, ${\bf R}(+, ;)$ has the decidable axiomatisation.
\end{theorem}

@@

\begin{corollary}
  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \in {\bf R}(;, +) \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
\end{corollary}

\begin{proof}
  Let $\mathcal{A}$ be a finite semilattice-ordered semigroup. We check axioms with the algorithm that yields ``no'', if $\mathcal{A}$ is not representable. So,
  In parallel, we check whether $\mathcal{A}$ has a finite representation. If there exists such $n < \omega$ such that $\mathcal{A}$ is representable over size $n$. Both sets are domains of corresponding computable functions, so they are enumerable. That makes ${\bf R}_{fin}(;, +)$ decidable by Post theorem.
\end{proof}
